---
title: "caf_diet"
author: "Yanireth Jimenez"
date: '2022-06-06'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
pacman::p_load(
  ggplot2,
  tidyverse,
  googlesheets4,
  ggpubr,
  ggrepel,
  lme4,
  lmerTest
)
```

# Load data

All relevant raw data is loaded here. read_sheet reads a csv file from 
google drive. Note that the csv file should be a google sheets file not .xlsx

```{r}
# 2h daily exposure to cafeteria diet
caf_diet_intake_raw <- read_sheet("https://docs.google.com/spreadsheets/d/1yN0byHvCX7_lhpqhMryK8NE8271o19uOHm12PpkQFE0/edit#gid=96738361")

# 22 hour chow (standard diet) intake
chow_diet_intake_raw <- read_sheet("https://docs.google.com/spreadsheets/d/1LhdjUG9xIxbZ3kmRWPWzSAUHyMce2JFTy2D1mtAGqCo/edit#gid=0")

# cafeteria diet macronutrient information
macronutrients_info <- read_sheet("https://docs.google.com/spreadsheets/u/1/d/1a7jy9vIBOiPlJYsKFF4vUyEtlqqNtbe9E5ALUA1d3l8/edit?usp=drive_web&ouid=104225426026972857358")

# weights
weights <- read_sheet("https://docs.google.com/spreadsheets/d/1OmcbGXrndiUeQcZvTu827imwuo2eFJEEOYs17NNugHw/edit#gid=993948139")
```

# Data preproc

```{r}
# fix variables data type
caf_diet_intake_raw <- caf_diet_intake_raw %>% 
  # set variables in gr as dbl
  mutate_at(
    vars(matches("gr")), as.numeric
  )
  
# daily cafeteria dieta intake
caf_diet_intake <- caf_diet_intake_raw %>% 
  # fix hour in and hour out format
  mutate(
    hour_in = hms::as_hms(hour_in),
    hour_out = hms::as_hms(hour_out),
    # corrected intake
    # criollitas
    criollitas_intake = criollitas_in_gr - (criollitas_out_gr + criollitas_sp_gr),
    # trencito
    trencito_intake = trencito_in_gr - (trencito_out_gr + trencito_sp_gr),
    # cheetos
    cheetos_intake = cheetos_in_gr - (cheetos_out_gr + cheetos_sp_gr),
    # papas
    papas_intake = papas_in_gr - (papas_out_gr + papas_sp_gr),
    # chow
    chow_intake = chow_in_gr - (chow_out_gr + chow_sp_gr)
  ) %>% 
  pivot_longer(
    cols = contains("intake"),
    names_to = "food_name",
    values_to = "corrected_intake_gr"
  ) %>% 
  mutate(
    food_name = str_extract(food_name, "[^_]*")
  )

# isolate caja control to obtain percentage increase/decrease of food
# without a mice
caja_control <- caf_diet_intake_raw %>% 
  filter(
    exp_tag == "caja_control"
  ) %>% 
  mutate(
    # fix hour format
    hour_in = hms::as_hms(hour_in),
    hour_out = hms::as_hms(hour_out),
    # criollitas
    criollitas_percent_diff = ((criollitas_out_gr + criollitas_sp_gr) - criollitas_in_gr) /
      criollitas_in_gr,
    # cheetos
    cheetos_percent_diff = ((cheetos_out_gr + cheetos_sp_gr) - cheetos_in_gr) /
      cheetos_in_gr,
    # trencito
    trencito_percent_diff = ((trencito_out_gr + trencito_sp_gr) - trencito_in_gr) /
      trencito_in_gr,
    # papas
    papas_percent_diff = ((papas_out_gr + papas_sp_gr) - papas_in_gr) /
      papas_in_gr,
    # chow
    chow_percent_diff = ((chow_out_gr + chow_sp_gr) - chow_in_gr) /
      chow_in_gr
  ) %>% 
  select(ID:hour_out, criollitas_percent_diff:chow_percent_diff) %>% 
  pivot_longer(
    cols = contains("percent_diff"),
    names_to = "food_name",
    values_to = "control_percent_diff"
  ) %>% 
  mutate(
    food_name = str_extract(food_name, "[^_]*")
  )

# add back caja control percent difference per food into caf_diet_intake
caf_diet_intake_cr <- caf_diet_intake %>% 
  left_join(
    caja_control %>% select(
      date,
      food_name,
      control_percent_diff
    ),
    by = c(
      "date",
      "food_name"
    )
  ) %>% 
  group_by(
    date,
    food_name,
    ID
  ) %>% 
  summarise(
    # corrected intake by spillage AND control cage
    # criollitas
    criollitas_intake_cr = (criollitas_in_gr + (criollitas_in_gr * control_percent_diff))
    - (criollitas_out_gr + criollitas_sp_gr),
    # trencito
    trencito_intake_cr = (trencito_in_gr + (trencito_in_gr * control_percent_diff))
    - (trencito_out_gr + trencito_sp_gr),
    # papas
    papas_intake_cr = (papas_in_gr + (papas_in_gr * control_percent_diff))
    - (papas_out_gr + papas_sp_gr),
    # cheetos
    cheetos_intake_cr = (cheetos_in_gr + (cheetos_in_gr * control_percent_diff))
    - (cheetos_out_gr + cheetos_sp_gr),
    # chow
    chow_intake_cr = (chow_in_gr + (chow_in_gr * control_percent_diff))
    - (chow_out_gr + chow_sp_gr)
  ) %>% 
    pivot_longer(
    cols = contains("intake_cr"),
    names_to = "food_name_cr",
    values_to = "corrected_intake_gr_cr"
  ) %>% 
  mutate(
    food_name = str_extract(food_name_cr, "[^_]*"),
    # if its less than 0.01 this cannot be measured by the scale
    corrected_intake_gr_cr = if_else(corrected_intake_gr_cr < 0.01, 0, corrected_intake_gr_cr)
  )

# macronutrient information
macronutrients_info_long <- macronutrients_info %>% 
  pivot_longer(
    cols = !macronutriente,
    names_to = "food_name",
    values_to = "kcal_per_gr"
  )

# 22 hour chow intake
chow_diet_intake <- chow_diet_intake_raw %>% 
  mutate(
    corrected_intake_gr = chow_22_in_gr - (chow_22_out_gr + chow_22_sp_gr),
    hour_in = hms::as_hms(hour_in),
    hour_out = hms::as_hms(hour_out)
  ) %>% 
  mutate(food_name = "chow", time = "22_hr")

chow_22_control <- chow_diet_intake_raw %>% 
  filter(
    exp_tag == "caja_control"
  ) %>% 
  mutate(
    hour_in = hms::as_hms(hour_in),
    hour_out = hms::as_hms(hour_out),
    # percent diff in control chow
    chow_22_percent_diff = ((chow_22_out_gr + chow_22_sp_gr) - chow_22_in_gr) /
      chow_22_in_gr,
  ) %>% 
  select(ID:hour_out, chow_22_percent_diff) %>% 
  pivot_longer(
    cols = contains("percent_diff"),
    names_to = "food_name",
    values_to = "control_percent_diff"
  ) %>% 
  mutate(
    food_name = str_extract(food_name, "[^_]*")
  )

chow_diet_intake_cr <- chow_diet_intake %>% 
  left_join(
    chow_22_control %>% select(
      date,
      food_name,
      control_percent_diff
    ),
    by = c(
      "date",
      "food_name"
    )
  ) %>% 
  group_by(
    date,
    food_name,
    ID
  ) %>% 
  summarise(
    # corrected intake by spillage AND control cage
    # chow
    chow_22_intake_cr = (chow_22_in_gr + (chow_22_in_gr * control_percent_diff))
    - (chow_22_out_gr + chow_22_sp_gr)
  ) %>% 
    pivot_longer(
    cols = contains("intake_cr"),
    names_to = "food_name_cr",
    values_to = "corrected_intake_gr_cr"
  ) %>% 
  mutate(
    food_name = str_extract(food_name_cr, "[^_]*"),
    # if its less than 0.01 this cannot be measured by the scale
    corrected_intake_gr_cr = if_else(corrected_intake_gr_cr < 0.01, 0, corrected_intake_gr_cr)
  )
```

## merge tables
```{r}
# this adds macro nutrients to each corrected food intake
caf_diet_intake_macros <- caf_diet_intake_cr %>% 
  left_join(
    macronutrients_info_long,
    by = c("food_name")
  ) %>% 
  mutate(time = "2_hr") %>% 
# next we compute the real kcal eaten for each food
  mutate(
    kcal_intake = kcal_per_gr * corrected_intake_gr_cr
  )

chow_22_intake_macros <- chow_diet_intake_cr %>% 
  left_join(
    macronutrients_info_long,
    by = c("food_name")
  ) %>% 
  mutate(
    kcal_intake = kcal_per_gr * corrected_intake_gr_cr
  )

# we bind both data sets to add up normal chow intake to the caf diet
# chow intake
caf_diet_intake_macros <- bind_rows(caf_diet_intake_macros, chow_22_intake_macros)
```

## summary tables
```{r}
# add back session info
session <- caf_diet_intake_raw %>% 
  select(
    ID,
    date,
    session,
    exp_tag
  )
caf_diet_intake_macros <- caf_diet_intake_macros %>% 
  left_join(
    session,
    by = c("ID", "date")
  )
# this tables considers total chow intake
# normal chow + chow consumed during cafeteria diet
caf_chow_kcal <- caf_diet_intake_macros %>% 
  group_by(
    ID,
    session,
    food_name,
    macronutriente,
    exp_tag
  ) %>% 
  summarise(
    kcal_intake = sum(kcal_intake)
  )

caf_chow_kcal$ID <- unlist(caf_chow_kcal$ID)

# normal + chow consumed during cafeteria diet vs caf diet in % of daily intake kcal
caf_chow_kcal_daily_ind <- caf_chow_kcal %>% 
  group_by(
    ID,
    session,
    exp_tag,
    macronutriente
  ) %>% 
  summarise(
    total_intake_sum_kcal = sum(kcal_intake)
  ) %>% 
  right_join(
    caf_chow_kcal,
    by = c("ID", "session", "exp_tag", "macronutriente")
  ) %>% 
  mutate(
    percent_of_daily_intake = (kcal_intake / total_intake_sum_kcal) * 100,
    type_of_food = if_else(food_name == "chow", "chow", "caf")
  ) %>% 
  group_by(
    ID,
    session,
    macronutriente,
    exp_tag,
    type_of_food
  ) %>% 
  summarise(
    percent_of_daily_intake_total = sum(percent_of_daily_intake)
  ) %>% 
  ungroup()
# this is the same data but at a group level
caf_chow_kcal_daily_group <- caf_chow_kcal_daily_ind %>% 
  filter(exp_tag != "caja_control", macronutriente == "total") %>% 
  drop_na() %>% 
  group_by(
    session,
    exp_tag,
    type_of_food
  ) %>% 
  summarise(
    percent_of_daily_intake_total_group = mean(percent_of_daily_intake_total),
    err = sd(percent_of_daily_intake_total) / sqrt(n())
  ) %>% 
  ungroup()

# intake by macro nutrient in kcal
# this considers only the 2hr intake
macro_intake_kcal_ind <- caf_diet_intake_macros %>%
  drop_na() %>% 
  filter(time == "2_hr") %>% 
  group_by(
    ID,
    session,
    macronutriente,
    exp_tag
  ) %>% 
  summarise(
    total_intake_kcal = sum(kcal_intake)
  ) 
macro_intake_kcal_group <- macro_intake_kcal_ind %>% 
  # at this point we have the total intake per session
  # now we get the mean and std error
  ungroup() %>% 
  group_by(
    session,
    macronutriente,
    exp_tag
  ) %>% 
  summarise(
    mean_total_intake_kcal = mean(total_intake_kcal),
    err = sd(total_intake_kcal) / sqrt(n())
  )

# intake in gr
# intake by macro nutrient in kcal
macro_intake_gr <- caf_diet_intake_macros %>% 
  drop_na() %>% 
  filter(time == "2_hr", macronutriente == "total") %>% 
  group_by(
    ID,
    session,
    exp_tag,
    food_name
  ) %>% 
  summarise(
    total_intake_gr = sum(corrected_intake_gr_cr)
  ) %>% 
  # at this point we have the total intake per session
  # now we get the mean and std error
  ungroup() %>% 
  group_by(
    session,
    exp_tag,
    food_name
  ) %>% 
  summarise(
    mean_total_intake_gr = mean(total_intake_gr),
    err = sd(total_intake_gr) / sqrt(n())
  )

# weight to consider change in weight relative to the start of 
# caf diet intake
weights_caf <- weights %>% 
  # session >= 1 are CAF sessions
  filter(session >= 1) %>% 
  group_by(
    ID
  ) %>% 
  # add group and difference in weight relative to the first session
  mutate(
    delta_weight = ((weight - weight[1]) / weight[1]) * 100,
    group = if_else(ID %in% c(338, 340, 342), "caf", "control")
  )

```


# Plots

## 2 hours kcal

x axis = experimental sessions
y axis = kcal or grams

```{r}
# macro intake in kcal
p1 <- macro_intake_kcal_ind %>% 
  mutate(ID = as.factor(ID)) %>% 
  filter(
    macronutriente == "total",
    exp_tag != "caja_control"
  ) %>% 
  ggplot(aes(
    session,
    total_intake_kcal,
    group = ID,
    color = exp_tag
  )) +
  geom_line(alpha = 0.2) +
  geom_point(alpha = 0.2) +
  geom_label_repel(aes(label = ID)) +
  geom_line(
    data = macro_intake_kcal_group %>% 
      filter(
        macronutriente == "total",
        exp_tag != "caja_control"
      ),
    inherit.aes = FALSE,
    aes(
      session,
      mean_total_intake_kcal,
      color = exp_tag
    )
  ) +
  geom_point(
      data = macro_intake_kcal_group %>% 
      filter(
        macronutriente == "total",
        exp_tag != "caja_control"
      ),
    inherit.aes = FALSE,
    aes(
      session,
      mean_total_intake_kcal,
      color = exp_tag
    )
  ) +
  geom_errorbar(
      data = macro_intake_kcal_group %>% 
      filter(
        macronutriente == "total",
        exp_tag != "caja_control"
      ),
    inherit.aes = FALSE,
    aes(
      session,
      mean_total_intake_kcal,
      color = exp_tag,
      ymin = mean_total_intake_kcal - err,
      ymax = mean_total_intake_kcal + err
    ),
    width = 0.01
  ) +
  theme_pubr() +
  labs(color = "Experimental condition") +
  ylab("Mean intake (kcal)") +
  xlab("Sessions") +
  scale_x_continuous(breaks = seq(1, max(macro_intake_kcal_ind$session), 1))
```

## 2 hours gr per food type
```{r}
# macro intake in gr
p2 <- macro_intake_gr %>% 
  filter(
    exp_tag != "caja_control"
  ) %>% 
  ggplot(aes(
    as.factor(session),
    group = food_name,
    mean_total_intake_gr,
    color = food_name,
    ymin = mean_total_intake_gr - err,
    ymax = mean_total_intake_gr + err
  )) +
  geom_point() +
  geom_line() +
  geom_errorbar(width = 0.1) +
  theme_pubr() +
  labs(color = "Experimental condition") +
  ylab("Mean intake (gr)") +
  xlab("Sessions") +
  facet_wrap(~exp_tag)
```


## percent intake in 24 hours of cafeteria dieta vs chow

x axis = experimental sessions
y axis = % of intake in kcal caf vs chow
```{r}
p3 <- caf_chow_kcal_daily_group %>% 
  filter(exp_tag != "caja_control") %>% 
  ggplot(aes(
    session,
    percent_of_daily_intake_total_group,
    color = type_of_food
  )) +
  geom_hline(yintercept = 50, color = "gray") +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(
    ymin = percent_of_daily_intake_total_group - err,
    ymax = percent_of_daily_intake_total_group + err
    ), width = 0.01) +
   geom_point(
     inherit.aes = FALSE,
    data = caf_chow_kcal_daily_ind %>% 
      filter(macronutriente == "total", exp_tag != "caja_control"),
    aes(
      session,
      percent_of_daily_intake_total,
      color = type_of_food
    ), alpha = 0.3
  ) +
  geom_label_repel(
    inherit.aes = FALSE,
    data = caf_chow_kcal_daily_ind %>% 
      filter(macronutriente == "total", exp_tag != "caja_control"),
    aes(
      session,
      percent_of_daily_intake_total,
      color = type_of_food,
      label = ID
    ), size = 2
  ) +
  facet_wrap(~exp_tag) +
  ylab("Mean percent of daily intake") +
  scale_x_continuous(breaks = seq(1, max(caf_chow_kcal_daily_group$session), 1)) +
  theme_pubr()
```

## weights
```{r}
# pick last labels
w_labs <- weights_caf %>% group_by(ID) %>% filter(session == max(session))
w1 <- weights_caf %>% 
  ggplot(aes(
    session,
    delta_weight,
    color = group,
    group = ID
  )) +
  geom_point() +
  geom_line() +
  geom_label_repel(aes(label = ID), data = w_labs) +
  geom_hline(yintercept = 0) +
  theme_pubr()

# group weight delta
wm <- weights_caf %>% 
  ungroup() %>% 
  group_by(
    group,
    session
  ) %>% 
  summarise(
    m = mean(delta_weight),
    s = sd(delta_weight) / sqrt(n())
  ) %>% 
  ungroup()

library(ggtext)
std_colors <- c("#000000", "#56B4E9", "#E69F00")
w1 <- wm %>% 
  ggplot(aes(
    session,
    m,
    ymin = m - s,
    ymax = m + s,
    group = group,
    color = group
  )) +
  geom_line(key_glyph = draw_key_blank) +
  geom_point(key_glyph = draw_key_blank) +
  geom_errorbar(key_glyph = draw_key_blank) +
  labs(colour = "") +
  scale_color_manual(labels = paste("<span style='color:",
                                   std_colors,
                                   "'>",
                                   c("Cafeteria", "Control"),
                                   "</span>"),
                    values = std_colors) +
  theme(legend.key = element_blank(),
        legend.text=element_markdown(size=12),
        legend.key.size = unit(0,"line"))


w2 <- weights_caf %>% 
  ggplot(aes(
    session,
    weight,
    color = group,
    group = ID
  )) +
  geom_point() +
  geom_line() +
  geom_label_repel(aes(label = ID), data = w_labs) +
  theme_pubr()
cowplot::plot_grid(w1, w2,
                   labels = c('percent change', 'weight'),
                   label_size = 12)
```


# save plots
```{r}
p1
p2
p3
```

# Statistical analysis

```{r}
data <- caf_chow_kcal %>% 
  filter(
    macronutriente == "total",
    exp_tag != "caja_control"
  ) %>% 
  mutate(
    food_cat = if_else(food_name == "chow", "chow", "caf")
  ) %>% 
  group_by(
    ID,
    session,
    exp_tag,
    food_cat
  ) %>% 
  summarise(
    kcal_intake_cat = sum(kcal_intake)
  ) %>% 
  ungroup()

mdl_intake <- lmer(
  data = data,
  kcal_intake_cat ~ exp_tag * food_cat + (1 | ID)
)
summary(mdl_intake)
emm <- emmeans::emmeans(mdl_intake, pairwise ~ exp_tag * food_cat)

# weight
data_caf <- weights_caf
mdl_weight <- lmer(
  data = data_caf,
  delta_weight ~ group * as.factor(session) + (1 | ID)
)
summary(mdl_weight)
emm_weight <- emmeans::emmeans(mdl_weight, pairwise ~ group | session)
```

